### 2023-09-14 main.tex
lkj

a particles of a give type (if present) is replaced with a

@@

\subsection{Integrable process on a line segment}
In this section we specialize the multi-species stirring process to the geometry of the one dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs act at the boundary sites $1$ and $L$. These boundaries, exchanging particle with the external environment, put the chain out of equilibrium. The Hamiltonian that we consider here is obtained by \eqref{OriginalHamiltonian} assuming that the conductances are
\begin{equation}
	\omega_{x,y}=\begin{cases}
		1 \quad \text{if}\quad |x-y|=1\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
and the local inhomogeneities are
\begin{equation}
	\Gamma_{x}=\begin{cases}
		1\quad \text{if} \quad x\in \{1,L\}\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
We further assume that the maximal occupancy in each site is $\nu=1$, i.e. that there is an \textit{hard-core interaction}.
We will denote the element of the representation of the $gl(N)$ Lie algebra by lowercase letters, i.e. $(e_{ab})_{a,b\in\{1,\ldots,N\}}$, to highlight that we are in  the fundamental representation of the algebra.
The Hamiltonian is
\begin{equation}\label{hamiltonian}
	H=H_{left}+H_{bulk}+H_{right}
\end{equation}
where
\begin{equation}
	H_{bulk}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}
\end{equation}
Here $\mathcal{H}_{x,x+1}$  is a copy (acting on the edge $x,x+1$) of the two-site Hamiltonian
\begin{equation}
	\begin{split}
		\mathcal{H}=P-\id
	\end{split}
\end{equation}
with 
\begin{equation}
	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}.
\end{equation} 
The action of $\mathcal{H}$ amounts to the permutation of the occupations the two sites, i.e.
\begin{equation}
	\begin{split}
		\mathcal{H}|n\rangle\otimes   |m\rangle&=|m\rangle \otimes |n\rangle-|n\rangle \otimes|m\rangle
	\end{split}
\end{equation}
The boundary terms are given by 
\begin{equation}
	H_{left}=\begin{pmatrix}
		\alpha_{1}-1&\alpha_{1}&\alpha_{1}&\ldots&\ldots&\alpha_{1}\\
		\alpha_{2}&\alpha_{2}-1&\alpha_{2}&\ldots&\ldots&\alpha_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\alpha_{N-1}&\alpha_{N-1}&\ldots&\ldots&\alpha_{N-1}-1&\alpha_{N-1}\\
		\alpha_{N}&\alpha_{N}&\ldots&\ldots&\alpha_{N}&\alpha_{N}-1
	\end{pmatrix}
\end{equation}
\begin{equation}
	H_{right}=\begin{pmatrix}
		\beta_{1}-1&\beta_{1}&\beta_{1}&\ldots&\ldots&\beta_{1}\\
		\beta_{2}&\beta_{2}-1&\beta_{2}&\ldots&\ldots&\beta_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\beta_{N-1}&\beta_{N-1}&\ldots&\ldots&\beta_{N-1}-1&\beta_{N-1}\\
		\beta_{N}&\beta_{N}&\ldots&\ldots&\beta_{N}&\beta_{N}-1
	\end{pmatrix}
\end{equation}
where, without loss of generality,  we assume that the parameters satisfy
\begin{equation}\label{ratesConditions}
	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1
\end{equation} 
We define also 
\begin{equation}\label{lambdaConditions}
	\lambda_{a}=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}\lambda_{a}=0\,.
\end{equation}
Under these assumption the Hamiltonian is integrable (see \cite{vanicat2017exact} for its derivation within the Quantum Inverse Scattering method).

@@

when the state spaces of the dual processes is finite, the generators and the duality function can be represented as with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 

@@

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=2}^{N}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{2,\ldots,N\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
\end{equation}
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{1}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{2,\ldots,N\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}
\end{equation}

@@

\paragraph{Proof of \eqref{edgeDualRealtion}.}To show this relation we need two 'ingredients'. First the existence of a similarity transformation between the Hamiltonian $\mathcal{H}_{x,y}$  and its transposed. As we will show, this similarity transformation is $R_{x}R_{y}$.  Second, the possibility of finding a symmetry for the edge Hamiltonian. Exploiting \eqref{hamiltonianCasimir}, we will search for a symmetry of the second Casimir $C_{2}$. As we will see, this symmetry is $\sum_{a=1}^{N-1}E_{a1}$ \\

@@

\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C_{2})$. Using the bilinearity of the coproduct operator and the fact that $C_{2}$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in gl(N)$ it holds
\begin{equation}
	[C_{2},E]=0\quad \Rightarrow\quad \left[\Delta (C_{2}),\Delta(E)) \right]=0
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{a1}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. Therefore, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. \\ Exploiting these considerations we write

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of the Hadamard formula with $B=\sum_{a=1}^{N-1}E_{a1}$ defined in \eqref{equationE} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b N}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b N}\exp{(E)}=E_{b N}
    \end{equation}
    because $E_{b 1}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b N}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=1}^{N-1}\left(E_{ab}\delta_{b N}-E_{b N}\delta_{ab}\right)=-E_{bN}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{Nb}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{N a}\exp{(E)}=E_{Na}+E_{NN}-\sum_{c=1}^{N-1}\left(E_{cN}+E_{ca}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

@@

d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}||\alpha^{x}|\sum_{a=1}^{N}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}\right)

@@

d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}||\alpha^{x}|\sum_{a=1}^{N-1}\left((a^{\dagger})_{a}^{u(x)}E_{Na}^{x}-E_{aa}^{x}\right)   

@@

In this setting the extra-sites are denoted by $0$ and $L+1$, since connected respectively with sites $1$ and $L$ of the chain. Moreover, the matrix $R_{x}$ defined in \eqref{Revmatrix} 

@@

	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{(E^{x})}\otimes \mathcal{D}_{L+1}

@@

@@

Using the above duality elements, we can compute the  $m$-point correlations between non-empty particles using the absorption probabilities. Consider a dual configuration $\hat{\xi}:=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$ where $a_{k}\in\{1,\ldots,N-1\}$ and $x_{k}\in\{1,\ldots,L\}$ for every $k\in\{1,\ldots,m\}$. Considering the elements \eqref{ElementsDualityMatrixChain} of the duality matrix computed in the dual configuration $\hat{\xi}$ we obtain that, for every $\bm{n}$

@@

therefore, its expectation with respect to the steady state non-equilibrium distribution gives the $m$-points correlations. Moreover, recalling that the dual process eventually voids the chain by putting all particles in the extra-sites $0$ and $L$ we have the following chain of equalities

@@

	\\&=\langle\hat{\bm{\xi}}\rangle_{stat,dual}=\sum_{t_{1}=0}^{1}\ldots \sum_{t_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{i_{k}}\beta_{a_{k}}^{1-i_{k}}\mathcal{P}(t_{1},\ldots,t_{m},\hat{\bm{\xi}})

@@

	\mathcal{P}(t_{1},\ldots,t_{m},\hat{\bm{\xi}}):=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{k=1}^{m}\left(t_{a}\delta_{0}+(1-t_{a})\delta_{L+1}\right)| \bm{\xi}(0)=\hat{\bm{\xi}}\right)
and they are called \textit{absorption probabilities}.

@@

@@

The MPA gives an abstract form of the steady state in terms of an algebra of operators. However, to write explicitly the steady state one should provide a representation of this algebra. Another approach is to directly use the commutators and compute the steady-state correlations using the commutation relations and the boundary actions. This path has been followed in \cite{vanicat2017exact} to compute one-point and two-point correlations. 
Here we follow the approach of working with the abstract commutation  relations. We will show that the duality property established in Section \ref{sectionDuality} simplifies substantially the  use of commutators in the actual

@@

The first transformation $S_{1}$ is closely related to the dual matrix. Moreover, we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalize the right boundary. The composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra, paying the price of having left boundary that depends on the difference between the densities of the two reservoirs.

The computation are organized as follows: we first define the similarity transformations $S_{1},S_{2}$ and the Hamiltonians $H^{'},H^{''}$. Second, we compute right eigenvector with vanishing eigenvalue of $H^{''}$ thanks to the simpler commutators. Finally, we invert the similarities and retrieve the right eigenvector with vanishing eigenvalue of $H^{'}$ and the steady state of $H$.

@@

By taking the product over the chain we also define

@@

	|\bm{\tau}\rangle =\begin{pmatrix}

@@

	\end{pmatrix}\qquad \forall a=2,\ldots N

>>>>

a particles of a given type (if present) is replaced with a

@@

\subsection{Integrable process on a line segment}\label{subsection-description-process-LINE}
In this section we specialize the multi-species stirring process to the geometry of the one dimensional chain with sites $\{1,\ldots,L\}$ where two reservoirs act at the boundary sites $1$ and $L$. These reservoirs, exchanging particle with the external environment, put the chain out of equilibrium. The Hamiltonian that we consider here is obtained by \eqref{OriginalHamiltonian} assuming that the conductances are
\begin{equation}
	\omega_{x,y}=\begin{cases}
		1 \quad \text{if}\quad |x-y|=1\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
and the local inhomogeneities are
\begin{equation}
	\Gamma_{x}=\begin{cases}
		1\quad \text{if} \quad x\in \{1,L\}\\
		0\quad \text{otherwise}
	\end{cases}
\end{equation}
We further assume that the maximal occupancy in each site is $\nu=1$, i.e. that there is an \textit{hard-core interaction}.
We will denote the element of the representation of the $gl(N)$ Lie algebra by lowercase letters, i.e. $(e_{ab})_{a,b\in\{1,\ldots,N\}}$, to highlight that we are in the first fundamental representation of the algebra.
The Hamiltonian is
\begin{equation}\label{hamiltonian}
	H=H_{left}+H_{bulk}+H_{right}
\end{equation}
where
\begin{equation}
	H_{bulk}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}
\end{equation}
Here $\mathcal{H}_{x,x+1}$  is a copy (acting on the edge $x,x+1$) of the two-site Hamiltonian
\begin{equation}\label{H-corsivo}
	\begin{split}
		\mathcal{H}=P-\id
	\end{split}
\end{equation}
with 
\begin{equation}
	P=\sum_{a,b=1}^Ne_{ab}\otimes e_{ba}.
\end{equation} 
The action of $\mathcal{H}$ amounts to the permutation of the occupations the two sites, i.e.
\begin{equation}
	\begin{split}
		\mathcal{H}|n\rangle\otimes   |m\rangle&=|m\rangle \otimes |n\rangle-|n\rangle \otimes|m\rangle
	\end{split}
\end{equation}
The boundary terms are given by 
\begin{equation}
	H_{left}=\begin{pmatrix}
		\alpha_{1}-1&\alpha_{1}&\alpha_{1}&\ldots&\ldots&\alpha_{1}\\
		\alpha_{2}&\alpha_{2}-1&\alpha_{2}&\ldots&\ldots&\alpha_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\alpha_{N-1}&\alpha_{N-1}&\ldots&\ldots&\alpha_{N-1}-1&\alpha_{N-1}\\
		\alpha_{N}&\alpha_{N}&\ldots&\ldots&\alpha_{N}&\alpha_{N}-1
	\end{pmatrix}
\end{equation}
\begin{equation}
	H_{right}=\begin{pmatrix}
		\beta_{1}-1&\beta_{1}&\beta_{1}&\ldots&\ldots&\beta_{1}\\
		\beta_{2}&\beta_{2}-1&\beta_{2}&\ldots&\ldots&\beta_{2}\\
		\vdots&\vdots& &\ddots& &\vdots\\
		\beta_{N-1}&\beta_{N-1}&\ldots&\ldots&\beta_{N-1}-1&\beta_{N-1}\\
		\beta_{N}&\beta_{N}&\ldots&\ldots&\beta_{N}&\beta_{N}-1
	\end{pmatrix}
\end{equation}
where, without loss of generality,  we assume that the parameters satisfy
\begin{equation}\label{ratesConditions}
	\sum_{a=1}^{N}\alpha_{a}=1,\qquad\sum_{a=1}^{N}\beta_{a}=1
\end{equation} 
We define also 
\begin{equation}\label{lambdaConditions}
	\lambda_{a}=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}\lambda_{a}=0\,.
\end{equation}
Under these assumption the Hamiltonian is integrable (see \cite{vanicat2017exact} for its derivation within the Quantum Inverse Scattering method).

@@

when the state spaces of the dual processes is finite, the generators and the duality function can be represented as matrices with elements $\mathcal{L}(\eta,\eta^{'})$, $\widetilde{\mathcal{L}}(\xi,\xi^{'})$ and $D(\eta,\xi)$ for arbitrary $\eta,\eta^{'}\in\Omega$ and $\xi,\xi^{'}\in \widetilde{\Omega}$. Therefore, we can write the duality relation \eqref{dualityRelationGenerator} as 

@@

The duality function is given by
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we denote the \textit{density} of the species $a\in \{1,\ldots,N-1\}$ imposed by the reservoir at site $x\in V$  by 
\begin{equation}
\rho_{a}^{x}:=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}
\end{equation}
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\textbf{Remark}: in the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigtimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}
\end{equation}

@@

\paragraph{Proof of \eqref{edgeDualRealtion}.}To show this relation we need two 'ingredients'. First the existence of a similarity transformation between the Hamiltonian $\mathcal{H}_{x,y}$  and its transposed. As we will show, this similarity transformation is $R_{x}R_{y}$.  Second, the possibility of finding a symmetry for the edge Hamiltonian. Exploiting \eqref{hamiltonianCasimir}, we will search for a symmetry of the second Casimir $C_{2}$. As we will see, this symmetry is $\sum_{a=1}^{N-1}E_{aN}$ \\

@@

\begin{equation}\label{transpositionPropertyH}
    \mathcal{H}_{x,y}^{T}=\left(R_{x}R_{y}\right)\mathcal{H}_{x,y}\left(R_{x}R_{y}\right)^{-1}
\end{equation}
therefore we have found the similarity transformation between $\mathcal{H}_{x,y}$ and its transposed. \\
We now look for a symmetry of $\mathcal{H}_{x,y}$, i.e. a matrix $S_{x,y}$ of the same dimension such that it satisfies 
\begin{equation}
	\mathcal{H}_{x,y}S_{x,y}=S_{x,y}\mathcal{H}_{x,y}\quad \Leftrightarrow \quad [\mathcal{H}_{x,y},S_{x,y}]=0
\end{equation}
Using \eqref{hamiltonianCasimir}, we observe that $\mathcal{H}_{x,y}$ is proportional to the coproduct of the second Casimir, up to a diagonal term. Therefore, it is enough to look for a symmetry of $\Delta (C_{2})$. Using the bilinearity of the coproduct operator and the fact that $C_{2}$ belongs to the centre of the Lie algebra, it is easy to show that for any matrix $E\in gl(N)$ it holds
\begin{equation}
	[C_{2},E]=0\quad \Rightarrow\quad \left[\Delta (C_{2}),\Delta(E)) \right]=0
\end{equation}
We chose 
\begin{equation}
	E=\sum_{a=1}^{N-1}E_{aN}
\end{equation}
and, for a fixed $(x,y)\in \mathcal{E}$, we define
\begin{equation}
	S_{x,y}=\exp{(\Delta^{x,y}(E))}=\exp{(E^{x})}\exp{(E^{y})}
\end{equation}
where $\Delta^{x,y}(E)$ is a copy of the coproduct acting on sites $x$ and $y$. Therefore, this operator $S_{x,y}$ satisfies
\begin{equation}\label{symmetryH}
	\left[S_{x,y},\mathcal{H}_{x,y}\right]=0
\end{equation}
i.e. it is a symmetry of $\mathcal{H}_{x,y}$. \\ Exploiting these considerations we write

@@

% However, they would lead to duality matrices where the number of dual particles is greater than the number of original ones, loosing the simplification of the dynamics of the dual process. 
 \begin{flushright}
     $\square$
 \end{flushright}
 \paragraph{Proof of \eqref{siteDualRelation}.} To prove \eqref{siteDualRelation} we need to transform via the Hadamard formula the transposed of the site Hamiltonian \eqref{siteDualHamiltonian} and then to  introduce properly a creation operator acting on an extra-site $u(x)$. % that we connect to every site $x\in V$ of the graph $G$. \\
 Considering $A,B\in gl(N)$, the Hadamard formula reads 
 \begin{equation}\label{HadamardFormula}
     \exp{(-B)}A\exp{(B)}=A-\left[B,A\right]+\frac{1}{2!}\left[B,\left[B,A\right]\right]-\frac{1}{3!}\left[B,\left[B,\left[B,A\right]\right]\right]+\ldots
 \end{equation}
In the following we need the application of the Hadamard formula with $B=\sum_{a=1}^{N-1}E_{aN}$ defined in \eqref{EquationEx} and with $A$ equal to some of the generator $E_{ab}$ of the Lie algebra. Therefore we compute 
\begin{enumerate}
    \item for $A=E_{b N}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_El1}
        \exp{(-E)}E_{b N}\exp{(E)}=E_{b N}
    \end{equation}
    because $E_{b N}$ commutes with $E$
    \item for $A=E_{bb}$ with $b \in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT_Ell}
        \exp{(-E)}E_{b b}\exp{(E)}=E_{b b}+E_{b N}
    \end{equation}
Indeed, using \eqref{eq:comgl} we have 
    \begin{equation}
       \left[E, E_{bb}\right]=\sum_{a=1}^{N-1}\left(E_{ab}\delta_{b N}-E_{b N}\delta_{ab}\right)=-E_{bN}
    \end{equation}
    Inserting the above commutator in \eqref{HadamardFormula}, we obtain \eqref{HT_Ell}.
    \item for $A=E_{Nb}$ with $b\in \{1,\ldots,N-1\}$, we obtain 
    \begin{equation}\label{HT-E1l}
        \exp{(-E)}E_{N a}\exp{(E)}=E_{Na}+E_{NN}-\sum_{c=1}^{N-1}\left(E_{cN}+E_{ca}\right)
    \end{equation}
  %\com{Why indices have names $b, k, j, a, b$? Fix one set for the generators}

@@

d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\,|\alpha^{x}|\sum_{a=1}^{N}\left(\rho_{a}^{x}E_{Na}^{x}-E_{aa}\right)

@@

d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\,|\alpha^{x}|\sum_{a=1}^{N-1}\left((a^{\dagger})_{a}^{u(x)}E_{Na}^{x}-E_{aa}^{x}\right)   

@@

Here we consider the chain defined in Subsection \ref{subsection-description-process-LINE}. In this setting the extra-sites are denoted by $0$ and $L+1$, connected respectively with sites $1$ and $L$ of the chain. Moreover, the matrix $R_{x}$ defined in \eqref{Revmatrix} 

@@

	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}

@@

where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
	\widetilde{H}_{left}=|\alpha|\sum_{a=1}^{N-1}\left((a^{\dagger})_{a}^{0}\,E_{Na}^{1}-E_{aa}^{1}\right)\qquad 	\widetilde{H}_{right}=|\alpha|\sum_{a=1}^{N-1}\left((a^{\dagger})_{a}^{L+1}\,E_{Na}^{L}-E_{aa}^{L}\right)

@@

Using the above duality elements, we can compute the  $m$-point correlations between non-empty particles using the so called absorption probabilities. Consider a dual configuration $\hat{\xi}:=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$ where $a_{k}\in\{1,\ldots,N-1\}$ and $x_{k}\in\{1,\ldots,L\}$ for every $k\in\{1,\ldots,m\}$. Considering an arbitrary $\bm{n}$, the elements \eqref{ElementsDualityMatrixChain} of the duality matrix computed in the dual configuration $\hat{\xi}$ read

@@

To compute the steady state $m$-point correlations of the process $\bm{n}(t)$ we recall that the dual process eventually voids the chain (by putting all particles in the extra-sites $0$ and $L$). Therefore, we have the following equalities

@@

	\\&=\langle\hat{\bm{\xi}}\rangle_{stat,dual}=\sum_{c_{1}=0}^{1}\ldots \sum_{c_{m}=0}^{1}\prod_{k=1}^{m}\alpha_{a_{k}}^{c_{k}}\beta_{a_{k}}^{1-c_{k}}\mathcal{P}(c_{1},\ldots,c_{m},\hat{\bm{\xi}})

@@

	\mathcal{P}(c_{1},\ldots,c_{m},\hat{\bm{\xi}}):=\mathbb{P}\left(\bm{\xi}(\infty)=\sum_{k=1}^{m}\left(c_{a}\delta_{0}+(1-c_{a})\delta_{L+1}\right)| \bm{\xi}(0)=\hat{\bm{\xi}}\right)
and they are called \textit{absorption probabilities}, namely the joint probability of having, for all $k=1,\ldots,m$, $c_{k}$ particles of species $a_{k}$ absorbed at left and $1-c_{k}$ absorbed at right. Therefore, by computing these absorption probabilities we determine the arbitrary non-equilibrium steady state correlations of the integrable multi-species stirring process.

@@

In this section we aim to use duality and the matrix product ansatz to determine the non equilibrium steady state and its arbitrary correlation for the integrable version of the multi-species stirring process, defined in Subsection \ref{subsection-description-process-LINE}. As it will turn out, duality simplifies the commutation relations of the matrix product ansatz allowing to determined a closed formula for all correlations. As a by-product, absorption probabilities can be explicitly determined, giving a further probabilistic interpretation for the non-equilibrium steady state. 

@@

The MPA gives an abstract form of the steady state in terms of an algebra of operators. However, to write explicitly the steady state one should provide a representation of this algebra. Another approach is to directly use the commutators and and the boundary actions compute the steady-state correlations. This path has been followed in \cite{vanicat2017exact} to compute one-point and two-point correlations. 
Here we follow the same approach. We will show that the duality property established in Section \ref{sectionDuality} simplifies substantially the  use of commutators in the actual

@@

The first transformation $S_{1}$ is closely related to the dual matrix. Moreover, we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalize the right boundary. The composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. \\

The computation are organized as follows: we first define the similarity transformations $S_{1},S_{2}$ and the Hamiltonians $H^{'},H^{''}$. Second, we compute right eigenvector with vanishing eigenvalue of $H^{''}$ thanks to the simpler commutators. Finally, we invert the similarities and retrieve the right eigenvector with vanishing eigenvalue of $H^{'}$ and  the arbitrary correlations and the absorption probabilities. The explicit expression of the steady state of $H$ is reported in appendix \ref{appB}.

@@

By taking the product over the chain

@@

	|\bm{\tau}\rangle=|\tau_{1},\ldots,\tau_{L}\rangle =\begin{pmatrix}

@@

	\end{pmatrix}\qquad \forall a=1,\ldots N-1
